Competing Bandits: \\ The Perils of Exploration under Competition

### 2019-02-07 paper/content/introduction.tex
tweaks
Minor edits in related work.

\xhdr{Machine learning.} Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions based on available information). MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background. We consider i.i.d. rewards, a well-studied and well-understood MAB model \cite{bandits-ucb1}. We focus on a well-known distinction between ``greedy" (exploitation-only) algorithms, ``naive" algorithms that separate exploration and exploitation, and ``smart" algorithms that combine them. Switching from ``greedy" to ``naive" to ``smart" algorithms involves substantial adoption costs in infrastructure and personnel training \cite{MWT-WhitePaper-2016,DS-arxiv}.

@@

\asedit{In ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}), an algorithm sets up a ``duel" between a pair of arms in each round, and only learns which arm has ``won". While this setting features a form of competition inside an MAB problem, it is very different from ours.}

The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system,
    \eg \cite{Kremer-JPE14,Frazier-ec14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
dynamic auctions
    (see \cite{DynAuctions-survey10} for background),
online ad auctions, \eg
    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm},
human computation
    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13},
and repeated actions, \eg
    \cite{Amin-auctions-nips13,Amin-auctions-nips14,Jieming-ec18}.

@@

\asedit{Whether data gives competitive advantage
has been studied theoretically in \cite{varian2018artificial,
  lambrecht2015can} and empirically in
\cite{bajari2018impact}.
While these papers find that small amounts
  of additional data do not provide significant improvement,
 % in training a more predictive model,
  they focus on learning in isolation.
  %, which overlooks certain aspects in a competition.
%\swdelete{These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.}
We ask how additional data helps when firms simultaneously solve a machine learning problem and compete for consumers. Due to the competition dynamics in our model, we find that even a small amount of additional data tends to help a lot.}

The first-mover advantage has been well-studied in other settings in economics and marketing, see survey \cite{kerin1992first}.

>>>>

\xhdr{Machine learning.} Multi-armed bandits (MAB) is a tractable abstraction for the tradeoff between exploration and \emph{exploitation} (making good near-term decisions). MAB problems have been studied for many decades, see \cite{Bubeck-survey12} for background. We consider i.i.d. rewards, a well-studied and well-understood MAB model \cite{bandits-ucb1}. We focus on a well-known distinction between ``greedy" (exploitation-only) algorithms, ``naive" algorithms that separate exploration and exploitation, and ``smart" algorithms that combine them. Switching from ``greedy" to ``naive" to ``smart" algorithms involves substantial adoption costs in infrastructure and personnel training \cite{MWT-WhitePaper-2016,DS-arxiv}.

@@

In ``dueling bandits" (e.g., \cite{Yue-dueling-icml09, Yue-dueling12}), an algorithm sets up a ``duel" between a pair of arms in each round, and only learns which arm has ``won". While this setting features a form of competition inside an MAB problem, it is very different from ours.

The interplay between exploration, exploitation and incentives has been studied in other scenarios: incentivizing exploration in a recommendation system,
    \eg \cite{Kremer-JPE14,Frazier-ec14,Che-13,ICexploration-ec15,Bimpikis-exploration-ms17},
dynamic auctions
    (see \cite{DynAuctions-survey10} for background),
online ad auctions, \eg
    \cite{MechMAB-ec09,DevanurK09,NSV08,Transform-ec10-jacm},
human computation
    \cite{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13},
and repeated auctions, \eg
    \cite{Amin-auctions-nips13,Amin-auctions-nips14,Jieming-ec18}.

@@

Whether data gives competitive advantage
has been studied theoretically in \cite{varian2018artificial,
  lambrecht2015can} and empirically in
\cite{bajari2018impact}.
While these papers find that small amounts
  of additional data do not provide significant improvement,
 % in training a more predictive model,
  they focus on learning in isolation.
  %, which overlooks certain aspects in a competition.
%\swdelete{These studies consider the effect of additional data in terms of helping towards solving a machine learning problem in isolation and argue that it does not play a large role.}
We ask how additional data helps when firms simultaneously solve a machine learning problem and compete for consumers. Due to the competition dynamics in our model, we find that even a small amount of additional data tends to help a lot. The first-mover advantage has been well-studied in other settings in economics and marketing, see survey \cite{kerin1992first}.
